Finiteness in Cubical Type Theory

### 2020-06-30 sections/finiteness-predicates.tex
down to surjections in split enumerability

This predicate is simple and useful, but we will see later on how it is perhaps
a little imprecise.
Before we dive in to exploring the predicate itself, though, we will need to
explain some of the terms we used in its definition.
\subsection{What is a List?}
We have referred to \(\AgdaDatatype{List}\) in the definition of split
enumerability.
In this paper, the type of lists is defined like so:
\begin{agdalisting} \label{list-def}
  \ExecuteMetaData[agda/Container/List.tex]{list-def}
\end{agdalisting}
This is the definition for a \emph{container} (Definition~\ref{container-def}):
effectively, the above definition says that ``Lists are a datatype whose shape
is given by the natural numbers, and which can be indexed by numbers smaller
than its shape''.

If that seems needlessly complex, don't worry: this definition is precisely
equivalent to the usual inductive one. \todo{This reads a little too informal I think}
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
\end{agdalisting}
And this isn't some kind of hand-waving equivalence, either: since we working in
HoTT, we can (and do) prove that the two types are equal, allowing us to use one
or the other depending on whichever is more convenient, and
\(\AgdaFunction{subst}\) in the other representation without loss of generality.
That said, defining lists as containers will reveal several interesting
connections and proofs about split enumerability and the other predicates, 
so for the remainder of the paper whenever we say \AgdaDatatype{List} we will
mean Equation~\ref{list-def}.

Before we get to those interesting proofs, though, there are some other things
that need defining.
\AgdaDatatype{Fin}, for instance: \(\AgdaDatatype{Fin}\;n\) is the type of
natural numbers smaller than \(n\).
Its definition relies on a type for disjoint union,
\(\AgdaDatatype{\ensuremath{\uplus}}\).
\begin{multicols}{2}
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/Fin/Base.tex]{fin-def}
  \end{agdalisting} \columnbreak
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}%
  % It is also expressible with only \(\Sigma\):
  % \begin{agdalisting}
  %   \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  % \end{agdalisting}
  % Although the inductive type definition is slightly more ergonomic.
\end{multicols} \noindent
And containers themselves, of course.
Containers are a well-studied topic in dependent type theory, with a rich
theory: we won't dive in to that here.
\begin{definition}[Containers] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S , P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{agdalisting} \label{container-interp}
    \ExecuteMetaData[agda/Container.tex]{container-interp}
  \end{agdalisting}
\end{definition}

@@

One of the nice things about containers is it gives us a generic way to define
``membership'':
\begin{multicols}{2}
  \begin{agdalisting} 
    \ExecuteMetaData[agda/Container/Membership.tex]{membership-def}
  \end{agdalisting} \columnbreak
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
  \end{agdalisting}
\end{multicols} \noindent 
Here we're using the homotopy-theory notion of a \AgdaDatatype{fiber} to define
membership: a fiber for some function \(f\) and some point \(y\) in its codomain
is a value \(x\) and a proof that \(f \; x \;\AgdaDatatype{≡}\;y\).
Membership also makes more sense when described concretely in terms of lists:
\(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) means ``there is an index
into \(\mathit{xs}\) such that the index points at an item equal to \(x\)''.
\subsection{Split Surjections}
We will now look at how this predicate relates to more traditional, classical
notions of finiteness.
In a classical setting we likely wouldn't mention ``lists'' or the like, and
would instead define finiteness based on the existence of some injection or
surjection.
As it turns out, our definition of finiteness here is precisely the same as the
surjection-based one, in quite a deep way!

First, we will need to define our terms: in HoTT, surjections are a little more
complex than what you'd find in either MLTT or classical mathematics.
\begin{definition}[Split Surjections] \label{split-surjections}
  We define \emph{split} surjections here \cite[definition
  4.6.1]{hottbook}.
  \begin{agdalisting} \label{sp-surj-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjective}
  \end{agdalisting} \vspace{-1.5\baselineskip}
  \begin{agdalisting} \label{sp-surj-arrow-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjection}
  \end{agdalisting}
\end{definition}
Over sets, the surjections and split surjections are the same thing, but there
is a difference one we involve non-set types like the circle.

We will now see that split enumerability is in fact a split surjection in
another form:
\begin{lemma} \label{split-enum-is-split-surj}%
  A proof of split enumerability is equivalent to a split surjection from a
  finite prefix of the natural numbers.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj-type}
  \end{agdalisting}
\end{lemma}
\begin{proof} \let\qed\relax \vspace{-1\baselineskip}
  \begin{minipage}[t]{.7\textwidth}\vspace{-\baselineskip}
    \begin{agdalisting*}
      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj}
    \end{agdalisting*}
  \end{minipage}
  \begin{minipage}[t]{.19\textwidth} \setstretch{1.1}

      Def.~\ref{split-enum-def} (\AgdaDatatype{\ensuremath{\mathcal{E}!}}) \\
      Eqn.~\ref{container-membership} (\AgdaDatatype{\ensuremath{\in}}) \\
      Eqn.~\ref{sp-surj-eqn}  \\
      Def.~\ref{List} (\(\AgdaDatatype{List}\)) \\
      Eqn.~\ref{container-interp}  \\
      Reassociation \\
      Eqn.~\ref{sp-surj-arrow-eqn}
  \end{minipage}
\end{proof}
In the above proof syntax the
\(\AgdaDatatype{\ensuremath{\equiv \langle{} \rangle{} }}\) connects lines which
are definitionally equal, i.e. they are ``obviously'' equal from the type
checker's perspective.
Clearly, only one line isn't a definitional equality: 
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/Sigma/Properties.tex]{reassoc}
\end{agdalisting}
(The simplicity of this proof, by the way, is why we preferred the
container-based definition of lists over the traditional one.)
\subsection{Instances} \todo{More intro to this}
We will now provide some example proofs of types which are split enumerable.
\begin{lemma}
  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
\end{lemma}
\begin{proof}
  These three types are quite obviously finite: we will show only the proof of
  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
  \begin{agdalisting} \label{bool-is-split-enumerable}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
  \end{agdalisting}
\end{proof}

@@

\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{agdalisting} \label{is-equiv-def}
    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{is-equiv-def}
  \end{agdalisting} \vspace{-.5\baselineskip}
  \begin{agdalisting} \label{equiv-def}
    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{equiv-def}
  \end{agdalisting}
\end{definition}
\begin{lemma} \label{bishop-equiv}
  Manifest bishop finiteness is equivalent to an equivalence to a finite prefix
  of the natural numbers.
  \begin{equation}
  \end{equation}
\end{lemma}
\begin{proof} \todo{Convert to Agda}
  \begin{align*}
     \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} }(\AgdaDatatype{\ensuremath{\mathcal{B}}})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
    && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\AgdaDatatype{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \simeq A )
    && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
  \end{align*}
\end{proof}
This proof is almost identical\footnotemark to the proof for
lemma~\ref{split-enum-is-split-surj}: it reveals that
enumeration-based finiteness predicates are simply another perspective on
relation-based ones.
\footnotetext{
  Unfortunately in our formalisation this proof cannot be a single line: for
  performance reasons \(\simeq\) is defined as a record type with eta-equality
  disabled, instead of the definition here which uses \(\Sigma\).
}

@@

And indeed this means that the circle is manifestly enumerable:
\begin{lemma}
  The circle \(S^1\) is manifestly enumerable.
\end{lemma}
\begin{proof}
  The support list firstly is a list containing the point constructor for the
  circle.
  Since the cover proof is truncated, we need only consider the point
  constructors of the circle: as such, the cover proof is essentially the same
  as the one for \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}(\top)\).
\end{proof}
\subsection{Surjections}
We already say that split enumerability was the listed form of a split
surjection: manifest enumerability is the listed form of a proper surjection.
\begin{definition}[Surjections] \label{surjections}
  We define proper surjections (not split surjections) here \cite[definition
  4.6.1]{hottbook}.
  \begin{alignat}{3}
    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn}
  \end{alignat}
\end{definition}
\begin{lemma}
  Manifest enumerability is equivalent to a surjection from a finite prefix of
  the natural numbers.
  \begin{equation}
    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\AgdaDatatype{Fin}\;n \twoheadrightarrow A)
  \end{equation}
\end{lemma}
\begin{proof}
  \begin{align*}
     \mathcal{E}(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
    && \text{def.~\ref{split-enum-def} }(\mathcal{E})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert \text{fib}_{\text{snd}(\mathit{xs})}(x) \rVert
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{surj-eqn} (surj)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{List} } (\AgdaDatatype{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{surj}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \twoheadrightarrow A )
    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow) \; \qedhere
  \end{align*}
\end{proof}
\subsection{Relation To Split Enumerability}
It is trivially easy to construct a proof that any split enumerable type is
manifest enumerable: we simply truncate the membership proof.
Going the other was requires us to extract a non-truncated proof from a
truncated one.
This proof relies on the following lemma:
\begin{lemma}
  We can ``recompute'' a truncated proof given a decision over a proof of the
  same type.
  \begin{equation}
    \frac{\lVert A \rVert \; \; \; \mathbf{Dec}(A)}{A}
  \end{equation}
\end{lemma}
\begin{proof}
  We proceed by case-analysis over the decision over \(A\).
  In the case where \(A\) is proven, we are done.
  In the case where \(A\) is disproven, we use lemma~\ref{prop-refute} to
  derive impossibility.
\end{proof}

>>>>

This predicate is simple and useful, but we will see later on how it is perhaps
a little imprecise.
Before we dive in to exploring the predicate itself, though, we will need to
explain some of the terms we used in its definition.
\subsection{What is a List?}
In this paper we prefer a slightly unusual definition for the type of lists:
\begin{agdalisting} \label{list-def}
  \ExecuteMetaData[agda/Container/List.tex]{list-def}
\end{agdalisting}
This is the definition for a \emph{container} (Definition~\ref{container-def}):
effectively, the above definition says that ``Lists are a datatype whose shape
is given by the natural numbers, and which can be indexed by numbers smaller
than its shape''.

\todo{The rest of this section certainly reads a lot better because of the
  inlined definitions, but I worry it's a little too informal and (possibly) too
verbose}

If that seems needlessly complex, don't worry: this definition is precisely
equivalent to the usual inductive one.
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/List/Base.tex]{list-def}
\end{agdalisting}
And this isn't some kind of hand-waving equivalence, either: since we working in
HoTT, we can (and do) prove that the two types are equal, allowing us to use one
or the other depending on whichever is more convenient, and
\(\AgdaFunction{subst}\) in the other representation without loss of generality.
That said, defining lists as containers will reveal several interesting
connections and proofs about split enumerability and the other predicates, 
so for the remainder of the paper whenever we say \AgdaDatatype{List} we will
mean Equation~\ref{list-def}.

Before we get to those interesting proofs, though, there are some other things
that need defining.
\AgdaDatatype{Fin}, for instance: \(\AgdaDatatype{Fin}\;n\) is the type of
natural numbers smaller than \(n\).
Its definition relies on a type for disjoint union,
\(\AgdaDatatype{\ensuremath{\uplus}}\).
\begin{multicols}{2}
  \begin{agdalisting}
    \ExecuteMetaData[agda/Data/Fin/Base.tex]{fin-def}
  \end{agdalisting} \columnbreak
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{disj-union}
  \end{agdalisting}%
  % It is also expressible with only \(\Sigma\):
  % \begin{agdalisting}
  %   \ExecuteMetaData[agda/Snippets/Introduction.tex]{sigma-disj-union}
  % \end{agdalisting}
  % Although the inductive type definition is slightly more ergonomic.
\end{multicols} \noindent
And containers themselves, of course.
Containers are a well-studied topic in dependent type theory, with a rich
theory: we won't dive in to that here.
\begin{definition}[Containers] \label{container-def}
  A container \cite{abbottContainersConstructingStrictly2005} is a pair
  \(S , P\) where \(S\) is a type, the elements of which are called
  the \emph{shapes} of the container, and \(P\) is a type family on \(S\), where
  the elements of \(P(s)\) are called the \emph{positions} of a container.
  We ``interpret'' a container into a functor defined like so:
  \begin{agdalisting} \label{container-interp}
    \ExecuteMetaData[agda/Container.tex]{container-interp}
  \end{agdalisting} \vspace{-\baselineskip}
\end{definition}

@@

One of the nice things about containers is it gives us a generic way to define
``membership'':
\begin{multicols}{2}
  \begin{agdalisting} \label{container-membership}
    \ExecuteMetaData[agda/Container/Membership.tex]{membership-def}
  \end{agdalisting} \columnbreak
  \begin{agdalisting}
    \ExecuteMetaData[agda/Snippets/Introduction.tex]{fiber}
  \end{agdalisting}
\end{multicols} \noindent 
Here we're using the homotopy-theory notion of a \AgdaDatatype{fiber} to define
membership: a fiber for some function \(f\) and some point \(y\) in its codomain
is a value \(x\) and a proof that \(f \; x \;\AgdaDatatype{≡}\;y\).
Membership also makes more sense when described concretely in terms of lists:
\(x\;\AgdaDatatype{\ensuremath{\in}}\;\mathit{xs}\) means ``there is an index
into \(\mathit{xs}\) such that the index points at an item equal to \(x\)''.
\subsection{Split Surjections}
Now that we have our terms defined, let's look a little at how split
enumerability relates to more traditional, classical notions of finiteness.
In a classical setting we likely wouldn't mention ``lists'' or the like, and
would instead define finiteness based on the existence of some injection or
surjection, say a surjection from a finite prefix of the natural numbers.
In HoTT, surjections (or, more precisely, \emph{split} surjections
\cite[definition 4.6.1]{hottbook}), are defined like so:
\begin{multicols}{2}
  \begin{agdalisting} \label{sp-surj-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjective}
  \end{agdalisting} \columnbreak
  \begin{agdalisting} \label{sp-surj-arrow-eqn}
    \ExecuteMetaData[agda/Function/Surjective/Base.tex]{split-surjection}
  \end{agdalisting}
\end{multicols}
As it turns out, our definition of finiteness here is precisely the same as a
surjection-based one, in quite a deep way!
\begin{lemma} \label{split-enum-is-split-surj}%
  A proof of split enumerability is equivalent to a split surjection from a
  finite prefix of the natural numbers.
  \begin{agdalisting}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj-type}
  \end{agdalisting}
\end{lemma}
\begin{proof} \let\qed\relax \vspace{-1\baselineskip}
  \begin{minipage}[t]{.7\textwidth}\vspace{-\baselineskip}
    \begin{agdalisting*}
      \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{is-split-inj}
    \end{agdalisting*}
  \end{minipage}
  \begin{minipage}[t]{.19\textwidth} \setstretch{1.1}

      Def.~\ref{split-enum-def} (\AgdaDatatype{\ensuremath{\mathcal{E}!}}) \\
      Eqn.~\ref{container-membership} (\AgdaDatatype{\ensuremath{\in}}) \\
      Eqn.~\ref{sp-surj-eqn}  \\
      Eqn.~\ref{list-def} (\(\AgdaDatatype{List}\)) \\
      Eqn.~\ref{container-interp}  \\
      Reassociation \\
      Eqn.~\ref{sp-surj-arrow-eqn}
  \end{minipage}
\end{proof}
In the above proof syntax the
\(\AgdaDatatype{\ensuremath{\equiv \langle{} \rangle{} }}\) connects lines which
are definitionally equal, i.e. they are ``obviously'' equal from the type
checker's perspective.
Clearly, only one line isn't a definitional equality: 
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/Sigma/Properties.tex]{reassoc}
\end{agdalisting}
This means that we could have in fact written the whole proof as follows:
\begin{agdalisting}
  \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{split-enum-is-split-surj-short}
\end{agdalisting}
The simplicity of this proof, by the way, is why we preferred the
container-based definition of lists over the traditional one.
\subsection{Instances} \todo{More intro to this}
We will now provide some example proofs of types which are split enumerable.
\begin{lemma}
  \(\bot\), \(\top\), and \(\AgdaDatatype{Bool}\) are split enumerable.
\end{lemma}
\begin{proof}
  These three types are quite obviously finite: we will show only the proof of
  finiteness for \(\AgdaDatatype{Bool}\) here for brevity's sake.
  \begin{agdalisting} \label{bool-is-split-enumerable}
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{bool-inst}
  \end{agdalisting}
\end{proof}

@@

\begin{definition}[Equivalences] \label{equivalences}
  We will take contractible maps \cite[definition 4.4.1]{hottbook} as our
  ``default'' definition of equivalences.
  \begin{agdalisting} \label{is-equiv-def}
    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{is-equiv-def}
  \end{agdalisting} \vspace{-.5\baselineskip}
  \begin{agdalisting} \label{equiv-def}
    \ExecuteMetaData[agda/Snippets/Equivalence.tex]{equiv-def}
  \end{agdalisting}
\end{definition}
\begin{lemma} \label{bishop-equiv}
  Manifest bishop finiteness is equivalent to an equivalence to a finite prefix
  of the natural numbers.
  \begin{equation}
  \end{equation}
\end{lemma}
\begin{proof} \todo{Convert to Agda}
  \begin{align*}
     \AgdaDatatype{\ensuremath{\mathcal{B}}}(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , x \inunique \mathit{xs}
    && \text{def.~\ref{bish-def} }(\AgdaDatatype{\ensuremath{\mathcal{B}}})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(x \in \mathit{xs})
    && \text{eqn.~\ref{uniq-memb-def} } (\inunique)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \text{isContr}(\text{fib}_{\text{snd}(\mathit{xs})}(x))
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{is-equiv-def} (isEquiv)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{isEquiv}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{isEquiv}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \simeq A )
    && \text{eqn.~\ref{equiv-def} } (\simeq) \; \qedhere
  \end{align*}
\end{proof}
This proof is almost identical\footnotemark to the proof for
lemma~\ref{split-enum-is-split-surj}: it reveals that
enumeration-based finiteness predicates are simply another perspective on
relation-based ones.
\footnotetext{
  Unfortunately in our formalisation this proof cannot be a single line: for
  performance reasons \(\simeq\) is defined as a record type with eta-equality
  disabled, instead of the definition here which uses \(\Sigma\).
}

@@

And indeed this means that the circle is manifestly enumerable:
\begin{lemma}
  The circle \(S^1\) is manifestly enumerable.
\end{lemma}
\begin{proof}
  The support list firstly is a list containing the point constructor for the
  circle.
  Since the cover proof is truncated, we need only consider the point
  constructors of the circle: as such, the cover proof is essentially the same
  as the one for \(\AgdaDatatype{\ensuremath{\mathcal{E}!}}(\top)\).
\end{proof}
\subsection{Surjections}
We already say that split enumerability was the listed form of a split
surjection: manifest enumerability is the listed form of a proper surjection.
\begin{definition}[Surjections] \label{surjections}
  We define proper surjections (not split surjections) here \cite[definition
  4.6.1]{hottbook}.
  \begin{alignat}{3}
    &\text{surj}(f)             &&\coloneqq \Pi(y : B) , \lVert \text{fib}_f(y) \rVert \label{surj-eqn} \\
    &A \twoheadrightarrow B     &&\coloneqq \Sigma (f : A \rightarrow B) , \text{surj}(f) \label{surj-arrow-eqn}
  \end{alignat}
\end{definition}
\begin{lemma}
  Manifest enumerability is equivalent to a surjection from a finite prefix of
  the natural numbers.
  \begin{equation}
    \mathcal{E}(A) \simeq \Sigma(n : \mathbb{N}) , (\AgdaDatatype{Fin}\;n \twoheadrightarrow A)
  \end{equation}
\end{lemma}
\begin{proof}
  \begin{align*}
     \mathcal{E}(A) &
    \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert x \in \mathit{xs} \rVert
    && \text{def.~\ref{split-enum-def} }(\mathcal{E})
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \Pi {(x : A)} , \lVert \text{fib}_{\text{snd}(\mathit{xs})}(x) \rVert
    && \text{eqn.~\ref{container-membership} } (\in)
    \\
    & \simeq \Sigma (\mathit{xs} : \textbf{List}(A)) , \text{surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{surj-eqn} (surj)}
    \\
    & \simeq \Sigma (\mathit{xs} : \llbracket \mathbb{N} , \AgdaDatatype{Fin} \rrbracket (A)) , \text{surj}(\text{snd}(\mathit{xs}))
    && \text{def.~\ref{list-def} } (\AgdaDatatype{List})
    \\
    & \simeq \Sigma (\mathit{xs} : \Sigma (n : \mathbb{N}) , \Pi (i : \AgdaDatatype{Fin}\;n) , A) , \text{surj}(\text{snd}(\mathit{xs}))
    && \text{eqn.~\ref{container-interp} } (\llbracket \wc \rrbracket)
    \\
    & \simeq \Sigma (n : \mathbb{N}) , \Sigma (f : \AgdaDatatype{Fin}\;n \rightarrow A) , \text{surj}(f)
    && \text{Reassociation of } \Sigma
    \\
    & \simeq \Sigma (n : \mathbb{N}) , ( \AgdaDatatype{Fin}\;n \twoheadrightarrow A )
    && \text{eqn.~\ref{surj-arrow-eqn} } (\twoheadrightarrow) \; \qedhere
  \end{align*}
\end{proof}
\subsection{Relation To Split Enumerability}
It is trivially easy to construct a proof that any split enumerable type is
manifest enumerable: we simply truncate the membership proof.
Going the other was requires us to extract a non-truncated proof from a
truncated one.
This proof relies on the following lemma:
\begin{lemma}
  We can ``recompute'' a truncated proof given a decision over a proof of the
  same type.
  \begin{equation}
    \frac{\lVert A \rVert \; \; \; \mathbf{Dec}(A)}{A}
  \end{equation}
\end{lemma}
\begin{proof}
  We proceed by case-analysis over the decision over \(A\).
  In the case where \(A\) is proven, we are done.
  In the case where \(A\) is disproven, we use lemma~\ref{prop-refute} to
  derive impossibility.
\end{proof}
